Practical implications of HATEOAS on a bridge between LOD and SPARQL

### 2012-02-27 eswc-2012/workshop.tex
Turtle that provides some links
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@122 43ef1190-90fd-4dd7-a769-d0121f02b80e
more concise in data set
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@123 43ef1190-90fd-4dd7-a769-d0121f02b80e
not pubby
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@124 43ef1190-90fd-4dd7-a769-d0121f02b80e
bashing graph store
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@126 43ef1190-90fd-4dd7-a769-d0121f02b80e

Semantic Web services is likely to be an important source of data for
future applications, but for foreseeable future just one of many. A
key promise of Linked Open Data (LOD) in particular, and the Semantic
Web in general is the ability to integrate many data sources
easily. However, this integration requires links and it requires that
the resulting graph can be traversed by reasonable means. While links
are quite abundant across the LOD cloud, this paper will argue that
key links are missing to make it possible to traverse the resulting
graph. Moreover, we will argue that this shortcoming is due to that
the community does not adequately take into account the constraint
known as ``Hypermedia as the Engine of Application State'' (commonly
abbreviated HATEOAS) from the REST architectural style, see
\cite{fielding}~Chapter~5.

@@

@@

\section{Finding the way with voiD}

@@

%\bibliographystyle{plain-csmin}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
%\bibliography{selectivity,federation,benchmarks,optimization}

>>>>

Semantic Web services is likely to be an important source of data for
future applications, but for foreseeable future just one of many. A
key promise of Linked Open Data (LOD) in particular, and the Semantic
Web in general is the ability to integrate many data sources
easily. However, this integration requires links and it requires that
the resulting graph can be traversed by reasonable means. While links
are quite abundant across the LOD cloud, this paper will argue that
key links are missing to make it possible to traverse the resulting
graph. Moreover, we will argue that this shortcoming is due to that
the community does not adequately take into account the constraint
known as ``Hypermedia as the Engine of Application State'' (commonly
abbreviated HATEOAS) from the REST architectural style, see
\cite{Fielding_2000_Architectural-Styles}~Chapter~5.

@@

Note that the bridge between LOD and SPARQL discussed in this paper is
different from the bridge provided by e.g. Pubby or RDF::LinkedData
and from the Linked Data API that is written to simplify use of linked
data by web developers\footnote{http://code.google.com/p/linked-data-api/}.

@@

\section{Lacking links in standards}

The SPARQL 1.1 Graph Store HTTP
Protocol\footnote{\url{http://www.w3.org/TR/2011/WD-sparql11-http-rdf-update-20110512/}},
currently with a Last Call status, claims to define a protocol in the
REST architectural style. It is not clear how it satisfies the
constraints of REST. On one hand, the strong ties to HTTP is overly
restrictive from a REST point of view, on the other, the specification
contains no mention today of how hypermedia (i.e. RDF in this case)
will drive interaction, which is the critical part of a using the REST
architectural style. Moreover, it defines out-of-band URI keywords
(\texttt{graph}). As mentioned in the introduction, this is not
necessarily a violation of architectural constraints, but the
specification has no mention of how applications can interact to
achieve the same tasks as it could with this out-of-band information.

This represents a lost opportunity to consider the practical
implications of HATEOAS to come up with a clearly RESTful specification.

% TODO: more than one URI for graph traversal

\section{Finding the way with VoID}

<> void:inDataset [ void:sparqlEndpoint </sparql> . ] .

@@

\bibliographystyle{plain-csmin}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
\bibliography{federation,webarch,webframeworks}
